title: "Gov 50 Final Project"
author: "Sally Edwards"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
    
##INTRODUCTION

In many of the past election cycles, immigration has taken a front-and-center position as one of the most salient topics of debate.  In broad, national terms, the debate has largely taken shape around geographic constraints.  This specific framing often portrays immigration as an issue which issue is practically concerning for voters who live close to the border, while citizens who live farther away are more likely to take a moralistic view of the situation. This implies that citizens who live in border cities will see the supposed negative ramifications of increased immigration in their everyday lives, and thus will be more prone to have a negative view of immigration. 

However, I would like to evaluate the extent of this truth -- as a Texan with family located in border cities and more interior towns, I can say from anecdotal evidence that this is not always the case. I believe that using data science to inform the conversation could help to develop a more nuanced understanding of both the issue and Texan voter's opinions. 

For my final project, I want to examine the relationship between opinion on immigration in Texas as compared to border town residence.  Simply put, does living in a border city in Texas make you more likely to have a negative view of immigration?  As border security has become an increasingly salient issue on the national political stage, it is all too easy for national politicians to make sweeping generalizations about the opinions of Texas voters without diving into the nuances of the multifaceted demographic. 

Immigration has become a central tenet of the GOP platform, and political figures — like Governor Greg Abbott or Senator Ted Cruz — have repeatedly framed the issue as immediately pressing and easily positionable.  However, I believe that individuals who live close to border towns will have a more nuanced view of this issue, possibly due to the fact that they have a higher personal relationship to the immigration process. My hypothesis is that individuals who are from border cities will view immigration as less of a concerning issue than the average Texan.


```{r}
library(tidyverse)
library(kableExtra)
library(sf)
library(ggspatial)
library(rnaturalearth)
library(maps)
library(rnaturalearthdata)

```

```{r}
citylist = c('El Paso, Texas')

city_data = tibble(location = citylist)

city_data <- city_data |>
  mutate(lat = 31.7619
         ,long = 106.4850) 

world_map_data <- ne_countries(scale = "medium", returnclass = "sf")
state_map_data<- map('state', fill = TRUE, plot = FALSE) %>% st_as_sf()

ggplot() +
  geom_sf(data = world_map_data) +
  geom_sf(data = state_map_data) +
  geom_point(data = city_data, aes(x = -106.4850, y = 31.7619), color = "navy", fill = "lavender" , pch=21, size=5, alpha=I(0.7))+
  coord_sf(xlim = c(-107, -94), ylim = c(26,37)) + 
  labs(title = "Map of Texas",
       x = "latitude",
       y = "longitude") 
```

I will primarily be examining this hypothesis through using the residents of the city of El Paso, TX, as a case study.  El Paso is one of Texas largest border cities, and is directly across the border from Juárez, Chihuahua, México.  In September of this year, El Paso's mayor Oscar Leeser said that the city has reached a "breaking point", as more than 2,000 migrants seek asylum in the city every day (https://www.reuters.com/world/americas/el-paso-texas-at-breaking-point-migrants-flood-border-mayor-says-2023-09-23/).  Thus, this provides a salient and immediate example of a border town which 

##Examining the Data
```{r}
library(haven)
library(kableExtra)
UT_data <- read_sav("utex0055d_output.sav")

UT_data_sample <- UT_data [1:5, c(1, 46, 161, 165, 166, 169)]

UT_data_sample|>
  kbl()|>
  kable_material(c("striped", "hover"))
```
##Data
In order to study the hypothesis, I plan to use data from August 2023 University of Texas/Texas Politics Project Poll.  My explanatory, independent variable will be an individual’s residential zip code (“zipcode”), collected from individual respondents, in the survey and the dependent variable will be an individual’s view of immigration.  This is quantified in the Q20_grid_A, wherein individuals rank how seriously the perceive the threat of immigration, from “extremely serious” (1) to "not serious" (5) along with options for “no opinion” or to skip.  If individuals with zip codes along the border rank their preferences from “not very serious” to “not serious” consistently while individuals in rural area rank it as a serious issue, then I believe my hypothesis will be proven.  However, if the opposite is true, such that individuals with zip codes along the border rank immigration as a serious issue (from 1-2) while individuals in rural area rank it as less serious — or if they align— then my hypothesis will be disproven.

Data Sourced from the Texas Politics Project and the University of Texas at Austin: August 2023 University of Texas/Texas Politics Project Poll https://texaspolitics.utexas.edu/polling-data-archive

Codebook -- paragraph references Q20 and “zipcode":
https://texaspolitics.utexas.edu/sites/texaspolitics.utexas.edu/files/utex0055d_codebook.pdf

Poll Summary: 
https://texaspolitics.utexas.edu/sites/texaspolitics.utexas.edu/files/202308_poll_topline.pdf
Poll Crosstabs:

(https://texaspolitics.utexas.edu/sites/texaspolitics.utexas.edu/files/202308_poll_crosstabs.pdf)

```{r}
library(tidyverse)
UT_data <- UT_data |>
  mutate(city = if_else(countyname == "El Paso TX", "El Paso", "Av. Texas")) |>
  mutate(is_El_Paso = if_else(city == "El Paso", 1, 0))

UT_data <- UT_data |>
  mutate(im_id = if_else(Q20_grid_A == `1`|`2`, 0, 1))


##Then, let's make a scatter plot to examine the relationship between immigration ideology and border town residence. 
EP_Scatter <- ggplot(data = UT_data, mapping = aes(x = city, y = Q20_grid_A)) +
                        geom_jitter(alpha = 0.75, color = "navy")+
  labs( x = "City",
        y = "How Seriously Immigration is Viewed as Threat",
        title = "Immigration Ideology in El Paso, TX") +
  theme_minimal()
EP_Scatter
```

##Final Project Results 

```{r}
UT_data |>
  ggplot(aes(x = factor(city), y = Q20_grid_A)) +
  geom_boxplot(color = "navy", fill = "lavender") + 
  labs(x = "City",
       y = "Ideology toward Immigration", 
       title = "Relationship Between Zip Code and Immigration Ideology")+ 
  theme_minimal()
```

```{r}
##difference in means of immigration ideology and border town residence 

imm_DIM<- UT_data |>
  group_by(city) |>
  summarize(avg_im_id = mean(Q20_grid_A, na.rm = TRUE)) |>
  pivot_wider(names_from = city, 
              values_from = avg_im_id) |>
  mutate(ATE = `El Paso` - `Av. Texas`) 

imm_DIM |>
  kbl()|>
  kable_material(c("striped", "hover"))
```

From this, we can see that there is a clear difference in immigration ideology depending on border town residence.  Resident in El Paso are more likely to have a warmer view towards immigration (closer to 1), as compared to the average Texan. 

From here, we must determine whether this difference is related to or caused by border town residence.  In order to do so, I will calculate the confidence interval for the above difference in means. 
```{r}
library(infer)
set.seed(02138)

ate_boots <- UT_data |>
  specify(im_id ~ is_El_Paso) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "diff in means")

ate_ci_95 <- ate_boots |>
  get_confidence_interval (level = 0.95)
ate_ci_95
```
Then, I will create a linear regression with the city of residence as the independent variable, and the immigration ideology as the dependent variable. 

```{r}
library(broom) 

imid_city_change <- lm(Q20_grid_A ~ city + birthyr + hispanic)

modelsummary :: modelsummary(
  imid_city_change, 
  statistic = c ("s.e. = {std.error}",
                 "p = {p.value}"),
  gof_map = c("nobs", "r.squared", "adj.r.squared"))
```